Outflows and rotation in Lyman alpha galaxies

### 2017-03-30 paper/paper.tex
section 2.1

@@

We model spherical galaxy to facilitate the interpretation of the results, by deleting what would be another degree of freedom regarding a different geometry. Furthermore, this approximation is commonly used in the literature, as it explains a wide variety of observational features (\cite{Ahn03}, \cite{Verhamme06}, \cite{Dijkstra06}). \\

There are 3 parameters in this model that define a LAE: the rotational velocity (vrot), the outflow velocity (vout) and the optical depth (tauh). vout is due to material ejected from the galaxy, by supernovas(\cite{Verhamme06}, \cite{Orsi12}, \cite{Hashimoto2015}, \cite{Gronke2015}). tauh roughly corresponds to number of H atoms found by a \lya photon if one traces a line from the center of the galaxy to its edge, and it resembles the mass of the LAE.\\

In this model, the LAE has a bulk velocity corresponding to the superposition of rotation and outflows, as shown in Fig. \ref{fig:model}. The velocity components are written in Eqs. (\ref{eq:vx}) (\ref{eq:vy}) (\ref{eq:vz}). In these equations, $R$ is the radius of the sphere; $x$, $y$ and $z$ are the coordinates in a cartesian frame; and the $\mp$ signs in $v_x$ and $v_y$ indicate the direction of rotation, respectively. This rotation is a solid body rotation and its direction goes according to the right hand rule applied to the $\hat{k}$ unit vector. The outflow velocity is dependent on the position relative to the center of the galaxy, being it zero at the center and maximum at the edge of the sphere.\\

\begin{equation}
	v_{x}=\frac{x}{R}v_{\rm out}-\frac{y}{R}v_{\rm rot} 
	\label{eq:vx}
\end{equation}

\begin{equation}
	v_{y}=\frac{y}{R}v_{\rm out}+\frac{x}{R}v_{\rm rot} 
	\label{eq:vy}
\end{equation}

\begin{equation}
	v_{z}=\frac{z}{R}v_{\rm out}
	\label{eq:vz}
\end{equation}

>>>>

CLARA follows the propagation of individual photons through a neutral
Hydrogen medium characterized by its temperature, velocity field and
global optical depth.
The code assumes an homogeneous density throughout the simulated
volume.
In the current implementation we neglect the influence of dust.
Our basic models is an spherical distribution of neutral hydrogen,
an approximation commonly used in the literature, as it explains a
wide variety of observational features \citep{Ahn03,Verhamme06,Dijkstra06}.

The central element in this paper is the velocity field that captures
outflows and rotation.
Outflows are captured by a Hubble-like radial velocity profile with
the velocity magnitude increasing linearly with the radial
coordinate; the outflow model is fully characterized by $V_{\rm out}$, the
velocity at the sphere's surface.
Rotation follows a solid body rotation profile, which is fully
characterized by $V_{\rm rot}$, the linear velocity at the sphere's surface.

The total velocity field corresponds to the superposition of rotation and
outflows.
The cartesian components take the form

\begin{equation}
	v_{x}=\frac{x}{R}V_{\rm out}-\frac{y}{R}V_{\rm rot} ,
	\label{eq:vx}
\end{equation}

\begin{equation}
	v_{y}=\frac{y}{R}V_{\rm out}+\frac{x}{R}V_{\rm rot} ,
	\label{eq:vy}
\end{equation}

\begin{equation}
	v_{z}=\frac{z}{R}V_{\rm out},
	\label{eq:vz}
\end{equation}
%
where $x$, $y$ and $z$ are the cartesian position coordinates with the
origin at the sphere's center $R$ is the radius of the sphere and the
direction of the angular velocity vector corresponds to the $\hat{k}$
unit vector.

For each model setup we follow $10^4$ individual photons generated at
the center of the sphere at the \lya\ line's center as they propagate
through the spere and finally scape.
We store the final frequency and propagation direction for each photon
at its last scattering.

@@
